% UTF-8
% Compile with XeLaTeX
% -----------------------------PAGE-----------------------------
\documentclass[12pt]{article}
\usepackage[a4paper,scale=0.8]{geometry}
\usepackage{xcolor}
% -----------------------------FONT-----------------------------
% Just to invoke fontspec, and also supports for CJK characters
\usepackage{xeCJK}
    \setCJKmainfont{SimSun}
    \setmainfont{Libertinus Serif}
% ---------------------------HYPERLINK---------------------------
\usepackage{hyperref}
    \hypersetup{
        colorlinks=true,
        allcolors=blue,
        pdfstartview=FitH,
        }
% ---------------------------PACKAGES---------------------------
\usepackage{metalogo}
\usepackage{gb4e}
% ========================DOCUMENT BEGINS========================
\begin{document}

\title{
    \href{https://youtu.be/yCnlaBJRKcE}{Молитва Франсуа Вийона}
    \\
    La prière de François Villon
    \\
    The Prayer of François Villon
    }
\author{Булат Шалвович Окуджава\\Bulat Shalvovich Okudzhava}
\date{}
\maketitle

\dotfill

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=2}{00:02}}   
    \glll    
    pɐˈka zʲɪˈmlʲa jɪˈɕːɵ ˈvʲertʲɪtsə pɐˈka jɪˈɕːɵ ˈjarək svʲet\\
    Пока Земля ещё вертится, пока ещё ярок свет.\\
    while Earth.\textsc{f.nom} still turn.3\textsc{sg.prs} while still bright.\textsc{m} light.\textsc{m.nom}\\
    \trans
    Tant que la terre tourne encore, tant que la lumière est vive.
    \trans
    While the earth is still turning, while the light is still bright.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=13}{00:13}}
    \glll
    ˈɡospədʲɪ daj ʐɨ tɨ ˈkaʐdəmʊ, tɕɪˈvo u nʲɪˈvo nʲet\\
    Господи, дай же Ты каждому, чего у него нет.\\
    Lord.\textsc{m.voc} give.2\textsc{sg.imp} then 2\textsc{sg} each.\textsc{m.dat} what.\textsc{gen} at 3\textsc{sg.m.gen} \textsc{neg}\\
    \trans
    Seigneur, donne à chacun ce qu'il n'a pas.
    \trans
    Lord, grant Thou to each man, that which he lacks.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=24}{00:24}}
    \glll
    ˈumnəmʊ daj ˈɡoɫəvʊ trʊsˈlʲivəmʊ daj kɐˈnʲa\\
    Умному дай голову, трусливому дай коня.\\
    clever.\textsc{m.dat} give.2\textsc{sg.imp} head.\textsc{f.acc} cowardly.\textsc{m.dat} give.2\textsc{sg.imp} horse.\textsc{m.acc}\\
    \trans
    Au sage une tête, au poltron un cheval.
    \trans
    To the wise man grant brains, to the coward a steed.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=34}{00:34}}
    \glll
    daj ɕːɪsˈlʲivəmʊ ˈdʲenʲɪk i nʲɪ zɐˈbutʲ prə mʲɪˈnʲa\\
    Дай счастливому денег, и не забудь про меня.\\
    give.2\textsc{sg.imp} happy.\textsc{m.dat} money.\textsc{pl.gen} and \textsc{neg} forget.2\textsc{sg.imp} about 1\textsc{sg.acc}\\
    \trans
    A l'heureux de l'argent, et ne m'oublie pas.
    \trans
    Grant the lucky man money, and don't forget about me.
\end{exe}

\dotfill

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=45}{00:45}}
    \glll
    pɐˈka zʲɪˈmlʲa jɪˈɕːɵ ˈvʲertʲɪtsə ˈɡospədʲɪ tvɐˈja vɫasʲtʲ\\
    Пока Земля ещё вертится, Господи, твоя власть.\\
    while Earth.\textsc{f.nom} still turn.3\textsc{sg.prs} Lord.\textsc{m.voc} your.\textsc{f.nom} authority.\textsc{f.nom}\\
    \trans
    Tant que la terre tourne encore, Seigneur, c'est en ton pouvoir !
    \trans
    While the earth is still turning, Lord, it is in Thy power!
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=56}{00:56}}
    \glll
    daj ˈrvuɕːɪmʊsʲə k ˈvɫasʲtʲɪ nɐˈvɫastvəvətsə fsɫasʲtʲ\\
    Дай рвущемуся к власти {навластвоваться} всласть.\\
    give.2\textsc{sg.imp} strive.\textsc{prs.ptcp.m.dat} to authority.\textsc{f.dat} \textcolor{red}{rule} to.one's.heart's.content\\
    \trans
    Donne à ceux qui veulent le pouvoir de régner à loisir.
    \trans
    Grant the man who wants power to rule to his heart's content.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=67}{01:07}}
    \glll
    daj pʲɪrʲɪˈdɨʂkʊ ˈɕːedrəmʊ xotʲ də ɪˈsxodə dʲnʲa\\
    Дай передышку щедрому, хоть до исхода дня.\\
    give.2\textsc{sg.imp} respite.\textsc{f.acc} generous.\textsc{m.dat} at.least until outcome.\textsc{m.gen} day.\textsc{m.gen}\\
    \trans
    Donne à souffler au généreux, au moins jusqu'au soir.
    \trans
    Grant the generous man a respite, if only to the end of the day.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=78}{01:18}}
    \glll
    ˈkaɪnʊ daj rɐˈskajɪnʲɪje i nʲɪ zɐˈbutʲ prə mʲɪˈnʲa\\
    Каину дай раскаяние, и не забудь про меня.\\
    Cain.\textsc{m.dat} give.2\textsc{sg.imp} repentance.\textsc{n.acc} and \textsc{neg} forget.2\textsc{sg.imp} about 1\textsc{sg.acc}\\
    \trans
    A Caïn donne le remords, et ne m'oublie pas.
    \trans
    To Cain grant repentence, and don't forget about me.
\end{exe}

\dotfill

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=88}{01:28}}
    \glll
    ja ˈznajʊ tɨ fsʲɵ ʊˈmʲejɪʂ ja ˈvʲerʊjʊ v ˈmudrəsʲtʲ tvɐˈju\\
    Я знаю, Ты всё умеешь, я верую в мудрость Твою.\\
    1\textsc{sg.nom} know.1\textsc{sg.prs} 2\textsc{sg.nom} all.\textsc{n.acc} be.able.to.2\textsc{sg.prs} 1\textsc{sg.nom} believe.1\textsc{sg.prs} in wisdom.\textsc{f.acc} your.\textsc{f.acc}\\
    \trans
    Je le sais : tu peux tout, je crois en ta sagesse.
    \trans
    All is in Thy power, I believe in Thy wisdom.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=100}{01:40}}
    \glll
    kak ˈvʲerʲɪt sɐɫˈdat ʊˈbʲitɨj ʂto on prəʐɨˈvajɪt v rɐˈju\\
    Как верит солдат убитый, что он проживает в раю.\\
    like believe.3\textsc{sg.prs} soldier.\textsc{m.nom} killed.\textsc{m.nom} \textsc{conj} 3\textsc{sg.m} live.3\textsc{sg.prs} in paradise.\textsc{m.loc}\\
    \trans
    Comme un soldat tué croit vivre en Paradis.
    \trans
    As the dead soldier believes he's living in heaven.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=110}{01:50}}
    \glll
    kak ˈvʲerʲɪt ˈkaʐdəjə ˈuxə ˈtʲixʲɪm rʲɪˈtɕam tvɐˈim\\
    как верит каждое ухо тихим речам твоим.\\
    like believe.3\textsc{sg.prs} each.\textsc{n.nom} ear.\textsc{n.nom} quiet.\textsc{pl.dat} speech.\textsc{pl.dat} your.\textsc{pl.dat}\\
    \trans
    Comme chaque oreille croit à tes doux propos.
    \trans
    As each ear believes Thy silent speeches.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=121}{02:01}}
    \glll
    kak ˈvʲerʊjɪm i mɨ ˈsamʲɪ nʲɪ ˈvʲedəjə ʂto tvɐˈrʲim\\
    Как веруем и мы сами, не ведая, что творим.\\
    like believe.1\textsc{pl.prs} and 1\textsc{pl.nom} self.\textsc{pl.nom} \textsc{neg} know.\textsc{prs.ptcp} what do.1\textsc{pl.prs}\\
    \trans
    Comme nous croyons nous-mêmes, ne sachant ce que nous faisons !
    \trans
    As we ourselves believe, not knowing what we do.
\end{exe}

\dotfill

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=133}{02:13}}
    \glll
    ˈɡospədʲɪ moj ˈboʐɨ zʲɪlʲɪnɐˈɡɫazɨj moj\\
    Господи мой Боже, зеленоглазый мой.\\
    Lord.\textsc{m.voc} my.\textsc{m.nom} God.\textsc{m.voc} green.eyed.\textsc{m.nom} my.\textsc{m.nom}\\
    \trans
    Seigneur, mon Dieu, mon doux Seigneur aux yeux verts !
    \trans
    Lord, my God, my green-eyed one!
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=144}{02:24}}
    \glll
    pɐˈka zʲɪˈmlʲa jɪˈɕːɵ ˈvʲertʲɪtsə i ˈɛtə jej ˈstranːə sɐˈmoj\\
    Пока Земля ещё вертится, и это ей странно самой.\\
    while Earth.\textsc{f.nom} still turn.3\textsc{sg.prs} and this 3\textsc{sg.f.dat} strange.\textsc{n} self.\textsc{f.dat}\\
    \trans
    Tant que tourne encore la terre, et cela paraît bien étrange.
    \trans
    While the earth is still turning, amazed it's still turning.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=155}{02:35}}
    \glll
    pɐˈka jɪˈɕːɵ xvɐˈtajɪt ˈvrʲemʲɪnʲɪ i ɐɡˈnʲa\\
    Пока ещё хватает, времени и огня.\\
    while still suffice.3\textsc{sg.prs} time.\textsc{n.gen} and fire.\textsc{m.gen}\\
    \trans
    Tant qu'il lui reste encore du temps et du feu.
    \trans
    While it still has time and fire.
\end{exe}

\begin{exe}
    \exi{\href{https://youtu.be/yCnlaBJRKcE?t=167}{02:47}}
    \glll
    daj ʐɨ tɨ fsʲem pənʲɪˈmnoɡʊ i nʲɪ zɐˈbutʲ prə mʲɪˈnʲa\\
    Дай же Ты всем понемногу, и не забудь про меня.\\
    give.2\textsc{sg.imp} then 2\textsc{sg} all.\textsc{pl.dat} little.by.little and \textsc{neg} forget.2\textsc{sg.imp} about 1\textsc{sg.acc}\\
    \trans
    Donne à chacun un peu, et ne m'oublie pas.
    \trans
    Grant Thou a little to everyone, and don't forget about me.
\end{exe}

\dotfill

\href{http://stengazeta.over-blog.com/article-boulat-okoudjava-fran-ois-villon-52672755.html}{Traduction} par Le chant du monde

\href{https://arlindo-correia.com/120505.html}{Translation} by Gerald Janecek

Glossed by Haoliang Zhao

Typeset by Haoliang Zhao using \XeLaTeX

\today

\end{document}
%=========================DOCUMENT ENDS=========================